Community-Centric Innovation

### 2014-02-20 paper.tex
cryptocurrencies

@@

>>>>

Beyond the purely technological aspects, the importance of community networks is precisely the fact that they are governed by a community.
Community governance is drastically different from hierarchical governance.
For the network to hold together, it is necessary that the participants reach some kind of consensus at the time of making decisions.
It is not uncommon to see different participants or groups of participants having completely different interests, ethical values or methodologies.
Still, it is in the benefit of all the participants to reach a consensus and agree in some basic principles.
The continuous negotiation, discussion and clashes of completely different positions can reach a level that threatens the integrity of the network.

The participation in a community network trains the citizens in self-management techniques that inherit many of the properties of decentralized networks.
The lack of a central authority obliges the participants to take responsibility, exercise their freedom respecting others and make concessions in order to achieve common goals.
The organizational structures that we find in community networks mimic those of free (libre) software communities and other Internet-based coordinated projects such as Wikipedia.

All participants are invited to contribute and contributions are evaluated in an informal process for merit or weaknesses.
The participants can progressively build up a reputation by making high-quality contributions and helping others to accomplish their goals.
The organization progressively evolves towards a meritocratic systems in which those participants that consistently contribute to the value of the community earn the respect of their peers and emerge as natural leaders.

Still, this is not a leadership which is attached to the person. 
The leadership is attached to the contributions.
This protects the community against individuals working for purely selfish interests that do not take into the account the goals of the community.

@@

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency.
The combination of cryptography, free libre software and a data network can result in networked cryptocurrencies.
In this technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules or protocol to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their proof-of-work and the majority decides which transactions are included in the public distributed ledger.

Cryptocurrency has some advantages when compared to traditional fiat currencies.
Some of the advantages of cryptocurrencies are related to their peer-to-peer nature.
One of them is the lack of a central authority that can make arbitrary decisions and create money at no cost.
The emission of money is controlled by the protocol and governed by publicly known rules.
This means that the availability of money at any point in the future is known by all the participants.
The emitted money is used to reward the nodes that contribute with hashing power to the security of the network.
Hashing power requires hardware and electricity, which means that is no longer possible to create money at zero cost.
In fact, it is a risky endeavour as the cost of hashing power can exceed the value of the generated money.

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
There is no risks of blockades of political interference.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency to bypass totalitarian governments blockades or lacking payment networks.

Some of the advantages of networked cryptocurrencies are derived from their networked nature.
The scope of the cryptocurrency is the same as the underlying network.
If a cryptocurrency runs on the Internet, the scope is global.
As a result, economic transactions benefit from all the advantages of Internet protocols.
Sending money is as simple as sending an electronic mail.
It propagates almost instantaneously across all the Internet and is confirmed in a matter of minutes.
Payments can be easily automatized using scripts or other software and emitted from any networked device, such as servers, laptops or mobile phones.
Furthermore, the network is available 24 a day 365 days a year.

Networked cryptocurrency communities level the playground for all participants.

As the source code for creating a
